Comparing Generative Pre-trained Transformer Models to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6939 Deaths in Sierra Leone from 2019-2022

### 2024-04-14 manuscript/manuscript.tex
Done child edits

Across the child age ranges, GPT had the highest PCCC at 0.69 to 0.75 for all age ranges (GPT-3.5 within 0.05 for 1-11 years), while the lowest PCCC values were for InSilicoVA for 1-11 years (InterVA-5 within 0.05 for 1-5 years), and InterVA-5 for 1-11 months (Figure \ref{fig-perf-agerange-child}). Each child age range had between 308 (12\%) to 1633 (62\%) records with most records in the 1-5 years range at 1633 records (62\%) and the other age ranges consisting of 12-14\% of all records. Figure \ref{fig-perf-cod-child} shows the highest and lowest performing models by PCCC across 9 child CODs. GPT-4 had the highest PCCC for 7 of 9 child CODs, where GPT-3.5 had 5, InterVA-5 had two (congenital anomalies and diarrhoeal diseases), and InSilicoVA had one (pneumonia) of the 7 CODs within 0.05 PCCC of GPT-4. InterVA-5 had the highest PCCC for diarrhoeal disease at 0.82 (GPT-4 at 0.81), and GPT-3.5 had the highest PCCC for ill-defined at 0.84. InSilicoVA had the lowest PCCC for 4 of 9 child CODs (InterVA-5 within 0.05 for 3 of the 4 CODs), InterVA-5 had the lowest PCC for the other 4 of 9 child CODs (InSilicoVA within 0.05 for injuries), and GPT-3.5 had the lowest PCCC for diarrhoeal diseases at 0.69. Congenital anomalies had one record only in which GPT-4, GPT-3.5, and InterVA-5 assigned a CGHR-10 code matching the physicians. Ill-defined had the largest difference between minimum and maximum PCCC at 0.84, while injuries, diarrhoeal disease, and pneumonia (with the exception of InterVA-5 as an outlier at 0.41) had higher PCCC values at or above 0.69 with small differences between minimum and maximum PCCC ranging from 0.01 to 0.14. Other infections had lower PCCC values at or below 0.54 with a small difference between minimum and maximum PCCC at 0.2. Each child COD had between 1 (\textless 1\%) to 1382 (52\%) records with malaria (n=1382, 52\%), other infections (n=667, 25\%), and diarrhoeal diseases (n=175, 7\%) making 84\% of all records. Congenital anomalies and nutritional deficiencies had less than 1\% of all records, while other CODs had between 1-3\% of all records. GPT-4 had the highest PCCC for both male and female records at 0.73 and 0.72 (GPT-3.5 at 0.69) respectively, while all other models had PCCC at or below 0.68 (see Figure \ref{fig-perf-sex-child} in Appendix \ref{app-perf-misc}).

>>>>

Across the child age ranges, the highest PCCC was observed for GPT-4 at 0.59 and 0.66 for 1-5 years and 6-11 years (GPT-3.5 within 0.05), GPT-3.5 at 0.54 for 6-11 months, and InSilicoVA at 0.5 for 1-5 months (Figure \ref{fig-perf-agerange-child}). Each child age range had between 308 (12\%) to 1633 (62\%) records with most records in the 1-5 years range at 1633 records (62\%) and the other age ranges consisting of 12-14\% of all records. For child CODs, the highest PCCC observed was for GPT-4 for 4 of 8 CODs, GPT-3.5 for 3 of 8 CODs (InSilicoVA within 0.05 for other infections), and InSilicoVA for pneumonia (Figure \ref{fig-perf-cod-child}). The lowest PCCC across child CODs was observed for InSilicoVA for epilepsy, leukaemia, and other noncommunicable diseases and malaria, InterVA-5 for pneumonia, nutritional deficiencies, and other infections. GPT-3.5, InSilicoVA, and InterVA-5 all had similar PCCC values greater than 0.8 as the lowest PCCC for injuries. Epilepsy, leukaemia, and other noncommunicable disease had the largest difference between minimum and maximum PCCC at 0.68, while Ill-defined had the next largest difference at 0.67 with InSilicoVA and InterVA-5 at 0 PCCC. Malaria and other infections had low PCCC values below 0.58 across models with small differences between minimum and maximum PCCC at 0.18 and 0.14 respectively. Injuries and diarrhoeal diseases had high PCC values above 0.6 across models with small differences between minimum and maximum PCC at 0.14 and 0.3 respectively. Each child COD had between 11 (\textless 1\%) to 1382 (52\%) records with malaria (n=1382, 52\%), other infections (n=667, 25\%), and diarrhoeal diseases (n=175, 7\%) making 84\% of all records. Nutritional deficiencies had less than 1\% of all records, while other CODs had between 1-3\% of all records. GPT-3.5 had the highest PCCC for both male and female records at 0.56 and 0.58 with GPT-4 within 0.05 PCCC at 0.55 and 0.54 respectively, while all other models had PCCC below 0.5 (see Figure \ref{fig-perf-sex-child} in Appendix \ref{app-perf-misc}).
